---
title: "Matching methods comparison"
date: "Compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

this <- system("hostname", TRUE)
if (this == "LAPTOP-IVSPBGCA") { 
	dpath <- "G:/.shortcut-targets-by-id/1mfeEftF_LgRcxOT98CBIaBbYN4ZHkBr_/share/image"
} else if (this == "DESKTOP-M2BA7AA") {
	dpath <- "google drive path"
}
```


```{r fileread, include=FALSE, cache=TRUE}
fdap <- list.files(file.path(dpath, "output/DAP"), pattern="xlsx$", recursive=TRUE, full=TRUE)
fdap <- fdap[!grepl("~", fdap)]
fdap <- fdap[!grepl("ETH", fdap)]

fibs <- gsub("DAP", "IBS", fdap)

ibs <- lapply(fibs, \(f) readxl::read_excel(f, "IBS_variety") |> data.frame())
dap <- lapply(fdap, \(f) {
    x <- readxl::read_excel(f, "res_full") |> data.frame()
    tab <- table(x$field_id[x$var_rank==1])
    exclude <- names(tab[tab>10])
    if (!is.null(exclude)) {
    	x <- x[!(x$field_id %in% exclude), ]
    }
    x
})

orders <- gsub("_IBS.xlsx", "", basename(fibs))

get1 <- function(x) {
  x <- x[x$var_rank == 1, ]
  x$var_rank <- NULL
  x
}
ibs1 <- lapply(ibs, get1)
dap1 <- lapply(dap, get1)

d <- lapply(1:length(dap), \(i) merge(ibs[[i]], dap[[i]], by=c("field_id", "ref_id")))
d1 <- lapply(1:length(dap), \(i) merge(ibs1[[i]], dap1[[i]], by="field_id"))
```

## Introduction

So far, we have two matching methods: IBS and DAP. Here we compare the results obtained with these methods. Right now we can compare the data for `r length(orders)` orders.

## Comparing the best matches

How many of the matches match?

```{r matchall}
tabs <- lapply(d1, \(x) table(x$variety.x == x$variety.y))
tabs <- do.call(rbind, tabs)
rownames(tabs) <- orders
knitr::kable(tabs, caption="(Dis-) agreement") |> kableExtra::kable_styling(full_width = FALSE)

```

How many of the matches match; only considering good matches (IBS > 0.7)?

```{r matchhigh}
tabs <- lapply(d1, \(x) { x <- x[x$IBS > .7, ]; 
                  table(x$variety.x == x$variety.y) })
tabs <- do.call(rbind, tabs)
rownames(tabs) <- orders
knitr::kable(tabs, caption="(Dis-) agreement") |> kableExtra::kable_styling(full_width = FALSE)
```

What is the rank in one method, for the best match in another method?

```{r pressure, echo=FALSE, fig.with=12}
rank_plot <- function(z, add=FALSE, main="") {
	M = c("IBS", "DAP")
	#z$var_rank[is.na(z$var_rank)] <- max(z$var_rank, na.rm=TRUE)
	z <- na.omit(z)
	if (add) {
		lines(sort(z$var_rank), (1:nrow(z))/nrow(z), lwd=2, col="red")	
		legend("bottomright", legend=paste(rev(M), "#1, ranked by", M), col=c("blue", "red"), lwd=2)
	} else {
		plot(sort(z$var_rank), (1:nrow(z))/nrow(z), xlab="Rank by other algorithm", 
			ylab="fraction of samples", las=1, type="l", lwd=2, col="blue", main=main)
	}
}
n <- length(dap1)
par(mfrow=c(1, n))
for (i in 1:n) {
   z1 <- merge(dap1[[i]], ibs[[i]], by=c("field_id", "variety"), all.x=TRUE)
   z2 <- merge(ibs1[[i]], dap[[i]], by=c("field_id", "variety"), all.x=TRUE)
   rank_plot(z1, main=gsub("_IBS.xlsx", "", orders[i]))
   rank_plot(z2, TRUE)
}

```




cor(d$IBS, d$Probability, use="pairwise.complete.obs")
cor(d$rank.x, d$rank.y, method = "spearman", use="pairwise.complete.obs")

dd <- d[d$rank.y <= 5 & d$rank.x <= 5, ]
cor(dd$IBS, dd$Probability, use="pairwise.complete.obs")
cor(dd$rank.x, dd$rank.y, method = "spearman", use="pairwise.complete.obs")

plot(d$IBS, d$Probability, cex=.1)

par(mfrow=c(1, 2))

i <- d$rank.x == 1
boxplot(d$rank.y[i])
j <- d$rank.y == 1
boxplot(d$rank.x[j])

i = d$rank.x < 5 | d$rank.y < 5
cor(d$IBS[i], d$Probability[i])

plot(d$rank.x[i], d$rank.y[i], cex=.1)


f = "C:/github/brian/IMAGE/data/ibs/Tanzania_Rice_matches.xlsx"
f = "C:/github/brian/IMAGE/data/ibs/Nigeria_Rice_matches.xlsx"

old <- readxl::read_excel(f, "IBS_cutoff_0.7_all_match")
old$field_id = gsub("DRi23-7955_", "", old$field_id)
old$ref_id = gsub("DRi23-7955_", "", old$ref_id)

dd <- merge(ib, old, by=c("field_id", "ref_id"))
plot(IBS.y ~ IBS.x, data=dd, cex=.1)




f = "c:/Users/rhijm/Downloads/res.csv"
x = read.csv(f)
a = x[,1:3]
b = x[,c(1, 4:5)]
m = merge(a, b, by=1:2)
plot(m[,3:4])
	
#}

f = "C:/github/brian/IMAGE/data/raw/snp/Tanzania/DArTPipelineData/rice_multi_match.csv"
do = read.csv(f)
i <- with(do, ave(Probability, SampleID, FUN=\(x) rank(1000 - x)))
do = do[i==1, ]
z <- merge(do, d, by=c("field_id", "ref_id"))


da1 = da1[order(da1[,1]), ]

d = d[order(d$field_Tid), ]
