---
title: "Matching methods comparison"
date: "Compiled on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
is_html <- knitr::is_html_output()
knitr::opts_chunk$set(echo=is_html)

this <- system("hostname", TRUE)
if (this == "LAPTOP-IVSPBGCA") { 
	dpath <- "G:/.shortcut-targets-by-id/1mfeEftF_LgRcxOT98CBIaBbYN4ZHkBr_/share/image"
} else if (this == "DESKTOP-M2BA7AA") {
	dpath <- "google drive path"
}

scrolltab <- function(data, caption) {
kableExtra::kbl(data, caption=caption) |>
    kableExtra::kable_classic(full_width = F,  position = "right") |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"))  |> 
    kableExtra::scroll_box(width = "100%", height = "400px")
}

```


```{r fileread, include=FALSE, cache=TRUE}
fdap <- list.files(file.path(dpath, "output/DAP"), pattern="xlsx$", recursive=TRUE, full=TRUE)
fdap <- fdap[!grepl("~", fdap)]
fibs <- gsub("DAP", "IBS", fdap)
ibs <- lapply(fibs, \(f) {
    d <- readxl::read_excel(f, "IBS_variety") |> data.frame()
    d <- d[d$IBS >= 0.8, ]
    d
})

dap <- lapply(fdap, \(f) {
    x <- readxl::read_excel(f, "res_full") |> data.frame()
    # exclude too many #1 calls
    tab <- table(x$field_id[x$var_rank==1])
    exclude <- names(tab[tab>10])
    if (!is.null(exclude)) {
    	x <- x[!(x$field_id %in% exclude), ]
    }
    x
})
#ham <- lapply(fham, \(f) readxl::read_excel(f, "HAM_variety") |> data.frame())

orders <- gsub("_IBS.xlsx", "", basename(fibs))
onms <- matchpoint:::order_names()
olabs <- onms$cc[match(orders, onms$name)]

get1 <- function(x) {
  x <- x[x$var_rank == 1, ]
  x$var_rank <- NULL
  x
}
ibs1 <- lapply(ibs, get1)
dap1 <- lapply(dap, get1)

d <- lapply(1:length(dap), \(i) {
    if (grepl("DEra22-7523", fdap[i])) {
       ibs[[i]]$ref_id <- gsub("_D.$", "", ibs[[i]]$ref_id)
    }
    merge(ibs[[i]], dap[[i]], by=c("field_id", "ref_id"))
})

d1 <- lapply(1:length(dap1), \(i) na.omit(merge(ibs1[[i]], dap1[[i]], by="field_id", all.x=T)))

GE <- ifelse(is_html, "≥", ">=")

pdftxt <- ifelse(is_html, '<a href="compare_methods.pdf">pdf version</a>', "")
```

`r pdftxt`

### Introduction

Comparing the results obtained with IBS and DAP. We are only considering IBS scores `r GE` 0.8.

### Best matches and correlation

What is the proportion of field samples to which both methods assign the same reference variety?

<div class = "row">
<div class = "col-md-6">
```{r match_high, fig.width=12, fig.height=5}

x <- sapply(d1, \(x) { c(sum(x$variety.x == x$variety.y) / nrow(x), nrow(x)) }) |> t() |> round(2)
y <- sapply(d1, \(x) { 
  x <- x[x$IBS > 0.9, ]
  c(sum(x$variety.x == x$variety.y) / nrow(x), nrow(x)) })  |> t() |> round(2)
colnames(x) <- colnames(y) <- c("agreement", "n-obs")
rownames(x) <- rownames(y) <- olabs

knitr::kable(x, caption=paste("IBS", GE, "0.8")) |> 
  kableExtra::kable_styling(full_width = FALSE)
```
</div>
<div class = "col-md-6">
```{r tab2}
knitr::kable(y, caption=paste("IBS", GE, "0.9")) |> 
  kableExtra::kable_styling(full_width = FALSE)
```
</div></div>


### Ranks

Given the best match of a field sample to a reference sample in one method; what is the rank of that reference sample in other methods (for IBS `r GE` 0.8)?

```{r pressure, fig.width=12, fig.height=5}
rank_plot <- function(z, add=FALSE, main="", ylab="") {
	M = c("IBS", "DAP")
	#z$var_rank[is.na(z$var_rank)] <- max(z$var_rank, na.rm=TRUE)
	z <- na.omit(z)
	if (add) {
		lines(sort(z$var_rank), (1:nrow(z))/nrow(z), lwd=2, col="red")
		legend("bottomright", legend=paste(rev(M), "#1, ranked by", M),
		       col=c("blue", "red"), lwd=c(4,2), cex=.75)
	} else {
		plot(sort(z$var_rank), (1:nrow(z))/nrow(z), 
		     xlab="Rank by other algorithm", 
			   ylab=ylab, axes=FALSE,
			   xlim=c(0, 10), yaxs="i", xaxs="i",
			   las=1, type="l", lwd=4, col="blue", main=main)
	   axis(1, 0:10)
	   axis(2, (0:10)/10, las=1)
	   grid(10,10)
	  grid()
	}
}
n <- length(dap1)
par(mfrow=c(2, 5), mar=c(4,4,2,0))

for (i in 1:n) {
   ylab <- ifelse(i %in% c(1,6), "fraction of samples", "")
   ib <- ibs[[i]]
   ib1 <- ibs1[[i]]
   z1 <- merge(dap1[[i]], ib, by=c("field_id", "variety"), all.x=TRUE)
   z2 <- merge(ib1, dap[[i]], by=c("field_id", "variety"), all.x=TRUE)
   rank_plot(z1, main=olabs[i], ylab=ylab)
   rank_plot(z2, TRUE)
}

```


### Variety counts 

Each dot is the number of calls for a reference variety by either method. The blue line is *y=x*. That is where the dots would be if there were no matching differences between the two methods. 

```{r besties, fig.width=12, fig.height=5}
m <- lapply(d1, \(x) {
  ibs <- as.data.frame(table(x$variety.x))
  dap <- as.data.frame(table(x$variety.y))
  m <- merge(ibs, dap, by=1, all=TRUE)
  m[is.na(m)] <- 0
  colnames(m)[2:3] <- c("IBS", "DAP")
  m
})

par(mfrow=c(2, 5), mar=c(4,4,2,0))
for (i in 1:length(m)) {
   left <- i %in% c(1,6)
   lim <- c(0, max(unlist(m[[i]][,-1])))
   plot(m[[i]][,-1], xlab="n (IBS)", ylab=ifelse(left, "n (DAP)", ""), 
        main=olabs[i], pch=20, col="red", las=1, cex=2, xpd=TRUE, xlim=lim, ylim=lim)
   grid()
   abline(0,1, col="blue")
}
```



### Distribution of best match scores 

For ease of visualization, DAP scores < 0.5 are not shown.

```{r match_scores, fig.width=12, fig.height=5}
par(mfrow=c(2, 5), mar=c(4,4,2,0))
for (i in 1:length(d1)) {
  y <- na.omit(d1[[i]])
  x <- (1:nrow(y))/nrow(y)

  left <- i %in% c(1,6)
  plot(x, sort(y$IBS), xlab="Fraction of samples", ylab=ifelse(left, "score", ""), las=1, ylim=c(0.5, 1), type="l", lwd=2, col="blue", main=olabs[i])
grid()
lines(x, sort(y$Probability), col="red", lwd=2)
if (i==1) {
  legend("topleft", legend=c("IBS", "DAP"), lwd=2, col=c("blue", "red"), cex=.9)
}
}
```

</br>

